---
title: "Chapter 3 - Random Forest Analysis"
output: html_document
---

# Import dataset:
```{r}
library(foreign)

setwd("/Users/avital/Documents/GitHub/dissertation/")
grid <- read.dbf("Aus_grid.dbf")
```


# Entire city model

## Split dataset
```{r}
set.seed(13)
train <- sample(nrow(grid),nrow(grid)*0.7)
training <- grid[train,]
validation <- grid[-train,]
```

## Create formulas
```{r}
formula <- as.vector(NA)

ind <- "Res + Com + Ins + Edu + Ind + Rec + Bikeway + BW_length + BW2roads + Bikeshare + Income + PopDensity + Bus + Bus_stops + Employment + IntDens + IntDensB + Entropy + Students + Car_pop + Males + Poverty + Whites + Distance"

for (i in 1:16){
  formula[i] <- paste0(colnames(grid[i+25])," ~ ",ind," + ",colnames(grid[41+(i-1)*3+1])," + ",colnames(grid[41+(i-1)*3+2])," + ",colnames(grid[41+(i-1)*3+3]))
}
```

## Create empty Increased MSE table
```{r}
IncMSE <- data.frame(Name=character(), R2=numeric(), MSE=numeric(), Trees=numeric(), Res=numeric(), Com=numeric(), Ins=numeric(), Edu=numeric(), Ind=numeric(), Rec=numeric(), Bikeway=numeric(), BW_length=numeric(), NW2roads=numeric(), Bikeshare=numeric(), Income=numeric(), PopDensity=numeric(), Bus=numeric(), Bus_stops=numeric(), Employment=numeric(), IntDens=numeric(), IntDensB=numeric(), Entropy=numeric(), Students=numeric(), Car_pop=numeric(), Males=numeric(), Poverty=numeric(), Whites=numeric(), Distance=numeric(), rho1K=numeric(), rho2K=numeric(), rho5K=numeric(), stringsAsFactors=FALSE)
```

## Create empty Increased Node Purity table
```{r}
IncNodePurity <- data.frame(Name=character(), Res=numeric(), Com=numeric(), Ins=numeric(), Edu=numeric(), Ind=numeric(), Rec=numeric(), Bikeway=numeric(), BW_length=numeric(), NW2roads=numeric(), Bikeshare=numeric(), Income=numeric(), PopDensity=numeric(), Bus=numeric(), Bus_stops=numeric(), Employment=numeric(), IntDens=numeric(), IntDensB=numeric(), Entropy=numeric(), Students=numeric(), Car_pop=numeric(), Males=numeric(), Poverty=numeric(), Whites=numeric(), Distance=numeric(), rho1K=numeric(), rho2K=numeric(), rho5K=numeric(), stringsAsFactors=FALSE)
```

## Figuring the best number of trees
```{r}
library(randomForest)

MS <- vector()
MS2 <- vector()
best <- vector()

set.seed(13)
for (i in 1:16){
  for (m in 1:50) {
    rf <- randomForest(formula=as.formula(formula[i]),data=training,importance=T,ntree=m*100) #examining ntrees between 100 to 5000
    yhat <- predict(rf, newdata=validation)
    MS[m] <- (yhat-validation[i+25])^2
    MS2[m] <- mean(MS[[m]])
  }
  best[i] <- which.min(MS2)*100
}
```

## Running RF models
```{r fig.height=6}
library(randomForest)
set.seed(13)
for (i in 1:16){
  rf <- randomForest(formula=as.formula(formula[i]),data=training,importance=T,ntree=best[i])
  yhat <- predict(rf, newdata=validation)
  MSE <- (yhat-validation[i+25])^2
  IncMSE[i,1] <- colnames(grid[i+25])
  IncNodePurity[i,1] <- colnames(grid[i+25])
  IncMSE[i,2] <- rf$rsq[length(rf$rsq)]
  IncMSE[i,3] <- mean(MSE[,1])
  IncMSE[i,4] <- best[i]
  imp <- importance(rf)
  for (j in 1:nrow(imp)){
    IncMSE[i,j+4] <- imp[j,1]
    IncNodePurity[i,j+1] <- imp[j,2]
  }
}
```


## Outputs

### Importance plot
```{r}
varImpPlot(rf)
```

### Increased MSE table
```{r}
library(knitr)
library(kableExtra)
library(dplyr)

t <- IncMSE %>% mutate_if(is.numeric, format, digits=2, nsmall=0) %>% t(.)
for (i in 1:ncol(t)){
  colnames(t)[i] <- t[1,i]
}
t <- t[-1,]

  t %>% kbl() %>% kable_paper(full_width=F) %>%
  column_spec (1:ncol(t),border_left = T, border_right = T, width="1cm") %>%
  kable_styling()
```

### Normalized increased MSE table
```{r}
IncMSE_norm <- IncMSE
range = vector()
IncMSE_norm[IncMSE_norm<0] <- 0

for (i in 5:ncol(IncMSE_norm)){
  min <- min(IncMSE_norm[,i])
  max <- max(IncMSE_norm[,i])
  range[i] <- max - min
  for (j in 1:nrow(IncMSE_norm)){
    IncMSE_norm[j,i] <- (IncMSE_norm[j,i]-min)/range[i]
  }
}

t <- IncMSE_norm %>% mutate_if(is.numeric, format, digits=2, nsmall=0) %>% t(.)
for (i in 1:ncol(t)){
  colnames(t)[i] <- t[1,i]
}
t <- t[-1,]

  t %>% kbl() %>% kable_paper(full_width=F) %>%
  column_spec (1:ncol(t),border_left = T, border_right = T, width="1cm") %>%
  kable_styling()
```

### Increased Node Purity
```{r}
library(knitr)
library(kableExtra)
library(dplyr)

t <- IncNodePurity %>% t(.)
for (i in 1:ncol(t)){
  colnames(t)[i] <- t[1,i]
}
t <- t[-1,]

  t %>% kbl() %>% kable_paper(full_width=F) %>%
  column_spec (1:ncol(t),border_left = T, border_right = T, width="1cm") %>%
  kable_styling()
```

### Normalized Increased Node Purity table
```{r}
IncMSE_norm <- IncNodePurity
range = vector()
IncMSE_norm[IncMSE_norm<0] <- 0

for (i in 2:ncol(IncMSE_norm)){
  min <- min(IncMSE_norm[,i])
  max <- max(IncMSE_norm[,i])
  range[i] <- max - min
  for (j in 1:nrow(IncMSE_norm)){
    IncMSE_norm[j,i] <- (IncMSE_norm[j,i]-min)/range[i]
  }
}

t <- IncMSE_norm %>% mutate_if(is.numeric, format, digits=2, nsmall=0) %>% t(.)
for (i in 1:ncol(t)){
  colnames(t)[i] <- t[1,i]
}
t <- t[-1,]

  t %>% kbl() %>% kable_paper(full_width=F) %>%
  column_spec (1:ncol(t),border_left = T, border_right = T, width="1cm") %>%
  kable_styling()
```


# City center model

```{r}
grid <- grid[which(grid$CityCenter==1),]
```

## Split dataset
```{r}
set.seed(13)
train <- sample(nrow(grid),nrow(grid)*0.7)
training <- grid[train,]
validation <- grid[-train,]
```

## Create formulas
```{r}
formula <- as.vector(NA)

ind <- "Res + Com + Ins + Edu + Ind + Rec + Bikeway + BW_length + BW2roads + Bikeshare + Income + PopDensity + Bus + Bus_stops + Employment + IntDens + IntDensB + Entropy + Students + Car_pop + Males + Poverty + Whites + Distance"

for (i in 1:16){
  formula[i] <- paste0(colnames(grid[i+25])," ~ ",ind," + ",colnames(grid[41+(i-1)*3+1])," + ",colnames(grid[41+(i-1)*3+2])," + ",colnames(grid[41+(i-1)*3+3]))
}
```

## Create empty Increased MSE table
```{r}
IncMSE <- data.frame(Name=character(), R2=numeric(), MSE=numeric(), Trees=numeric(), Res=numeric(), Com=numeric(), Ins=numeric(), Edu=numeric(), Ind=numeric(), Rec=numeric(), Bikeway=numeric(), BW_length=numeric(), NW2roads=numeric(), Bikeshare=numeric(), Income=numeric(), PopDensity=numeric(), Bus=numeric(), Bus_stops=numeric(), Employment=numeric(), IntDens=numeric(), IntDensB=numeric(), Entropy=numeric(), Students=numeric(), Car_pop=numeric(), Males=numeric(), Poverty=numeric(), Whites=numeric(), Distance=numeric(), rho1K=numeric(), rho2K=numeric(), rho5K=numeric(), stringsAsFactors=FALSE)
```

## Create empty Increased Node Purity table
```{r}
IncNodePurity <- data.frame(Name=character(), Res=numeric(), Com=numeric(), Ins=numeric(), Edu=numeric(), Ind=numeric(), Rec=numeric(), Bikeway=numeric(), BW_length=numeric(), NW2roads=numeric(), Bikeshare=numeric(), Income=numeric(), PopDensity=numeric(), Bus=numeric(), Bus_stops=numeric(), Employment=numeric(), IntDens=numeric(), IntDensB=numeric(), Entropy=numeric(), Students=numeric(), Car_pop=numeric(), Males=numeric(), Poverty=numeric(), Whites=numeric(), Distance=numeric(), rho1K=numeric(), rho2K=numeric(), rho5K=numeric(), stringsAsFactors=FALSE)
```

## Figuring the best number of trees
```{r}
library(randomForest)

MS <- vector()
MS2 <- vector()
best <- vector()

set.seed(13)
for (i in 1:16){
  for (m in 1:50) {
    rf <- randomForest(formula=as.formula(formula[i]),data=training,importance=T,ntree=m*100) #examining ntrees between 100 to 5000
    yhat <- predict(rf, newdata=validation)
    MS[m] <- (yhat-validation[i+25])^2
    MS2[m] <- mean(MS[[m]])
  }
  best[i] <- which.min(MS2)*100
}
```

## Running RF models
```{r fig.height=6}
library(randomForest)
set.seed(13)
for (i in 1:16){
  rf <- randomForest(formula=as.formula(formula[i]),data=training,importance=T,ntree=best[i])
  yhat <- predict(rf, newdata=validation)
  MSE <- (yhat-validation[i+25])^2
  IncMSE[i,1] <- colnames(grid[i+25])
  IncNodePurity[i,1] <- colnames(grid[i+25])
  IncMSE[i,2] <- rf$rsq[length(rf$rsq)]
  IncMSE[i,3] <- mean(MSE[,1])
  IncMSE[i,4] <- best[i]
  imp <- importance(rf)
  for (j in 1:nrow(imp)){
    IncMSE[i,j+4] <- imp[j,1]
    IncNodePurity[i,j+1] <- imp[j,2]
  }
}
```


## Outputs

### Importance plot
```{r}
varImpPlot(rf)
```

### Increased MSE table
```{r}
library(knitr)
library(kableExtra)
library(dplyr)

t <- IncMSE %>% mutate_if(is.numeric, format, digits=2, nsmall=0) %>% t(.)
for (i in 1:ncol(t)){
  colnames(t)[i] <- t[1,i]
}
t <- t[-1,]

  t %>% kbl() %>% kable_paper(full_width=F) %>%
  column_spec (1:ncol(t),border_left = T, border_right = T, width="1cm") %>%
  kable_styling()
```

### Normalized increased MSE table
```{r}
IncMSE_norm <- IncMSE
range = vector()
IncMSE_norm[IncMSE_norm<0] <- 0

for (i in 5:ncol(IncMSE_norm)){
  min <- min(IncMSE_norm[,i])
  max <- max(IncMSE_norm[,i])
  range[i] <- max - min
  for (j in 1:nrow(IncMSE_norm)){
    IncMSE_norm[j,i] <- (IncMSE_norm[j,i]-min)/range[i]
  }
}

t <- IncMSE_norm %>% mutate_if(is.numeric, format, digits=2, nsmall=0) %>% t(.)
for (i in 1:ncol(t)){
  colnames(t)[i] <- t[1,i]
}
t <- t[-1,]

  t %>% kbl() %>% kable_paper(full_width=F) %>%
  column_spec (1:ncol(t),border_left = T, border_right = T, width="1cm") %>%
  kable_styling()
```

### Increased Node Purity
```{r}
library(knitr)
library(kableExtra)
library(dplyr)

t <- IncNodePurity %>% t(.)
for (i in 1:ncol(t)){
  colnames(t)[i] <- t[1,i]
}
t <- t[-1,]

  t %>% kbl() %>% kable_paper(full_width=F) %>%
  column_spec (1:ncol(t),border_left = T, border_right = T, width="1cm") %>%
  kable_styling()
```

### Normalized Increased Node Purity table
```{r}
IncMSE_norm <- IncNodePurity
range = vector()
IncMSE_norm[IncMSE_norm<0] <- 0

for (i in 2:ncol(IncMSE_norm)){
  min <- min(IncMSE_norm[,i])
  max <- max(IncMSE_norm[,i])
  range[i] <- max - min
  for (j in 1:nrow(IncMSE_norm)){
    IncMSE_norm[j,i] <- (IncMSE_norm[j,i]-min)/range[i]
  }
}

t <- IncMSE_norm %>% mutate_if(is.numeric, format, digits=2, nsmall=0) %>% t(.)
for (i in 1:ncol(t)){
  colnames(t)[i] <- t[1,i]
}
t <- t[-1,]

  t %>% kbl() %>% kable_paper(full_width=F) %>%
  column_spec (1:ncol(t),border_left = T, border_right = T, width="1cm") %>%
  kable_styling()
```


# Bikesharing model

```{r}
grid <- read.dbf("Aus_SA.dbf")
```

## Split dataset
```{r}
set.seed(13)
train <- sample(nrow(grid),nrow(grid)*0.7)
training <- grid[train,]
validation <- grid[-train,]
```

## Create formulas
```{r}
formula <- as.vector(NA)

ind <- "Res + Com + Ins + Edu + Ind + Rec + Bikeway + BW_length + BW2roads + Income + PopDensity + Bus + Bus_stops + Employment + IntDens + IntDensB + Entropy + Students + Car_pop + Males + Poverty + Whites + Distance"

for (i in 1:8){
  formula[i] <- paste0(colnames(grid[i+25])," ~ ",ind," + ",colnames(grid[i+49])," + ",colnames(grid[i+41])," + ",colnames(grid[i+33]))
}
```

## Create empty Increased MSE table
```{r}
IncMSE <- data.frame(Name=character(), R2=numeric(), MSE=numeric(), Trees=numeric(), Res=numeric(), Com=numeric(), Ins=numeric(), Edu=numeric(), Ind=numeric(), Rec=numeric(), Bikeway=numeric(), BW_length=numeric(), NW2roads=numeric(), Income=numeric(), PopDensity=numeric(), Bus=numeric(), Bus_stops=numeric(), Employment=numeric(), IntDens=numeric(), IntDensB=numeric(), Entropy=numeric(), Students=numeric(), Car_pop=numeric(), Males=numeric(), Poverty=numeric(), Whites=numeric(), Distance=numeric(), rho1K=numeric(), rho2K=numeric(), rho5K=numeric(), stringsAsFactors=FALSE)
```

## Create empty Increased Node Purity table
```{r}
IncNodePurity <- data.frame(Name=character(), Res=numeric(), Com=numeric(), Ins=numeric(), Edu=numeric(), Ind=numeric(), Rec=numeric(), Bikeway=numeric(), BW_length=numeric(), NW2roads=numeric(), Income=numeric(), PopDensity=numeric(), Bus=numeric(), Bus_stops=numeric(), Employment=numeric(), IntDens=numeric(), IntDensB=numeric(), Entropy=numeric(), Students=numeric(), Car_pop=numeric(), Males=numeric(), Poverty=numeric(), Whites=numeric(), Distance=numeric(), rho1K=numeric(), rho2K=numeric(), rho5K=numeric(), stringsAsFactors=FALSE)
```

## Figuring the best number of trees
```{r}
library(randomForest)

MS <- vector()
MS2 <- vector()
best <- vector()

set.seed(13)
for (i in 1:8){
  for (m in 1:50) {
    rf <- randomForest(formula=as.formula(formula[i]),data=training,importance=T,ntree=m*100) #examining ntrees between 100 to 5000
    yhat <- predict(rf, newdata=validation)
    MS[m] <- (yhat-validation[i+25])^2
    MS2[m] <- mean(MS[[m]])
  }
  best[i] <- which.min(MS2)*100
}
```

## Running RF models
```{r fig.height=6}
library(randomForest)
set.seed(13)
for (i in 1:4){
  rf <- randomForest(formula=as.formula(formula[i]),data=training,importance=T,ntree=best[i])
  yhat <- predict(rf, newdata=validation)
  MSE <- (yhat-validation[i+25])^2
  IncMSE[i,1] <- colnames(grid[i+25])
  IncNodePurity[i,1] <- colnames(grid[i+25])
  IncMSE[i,2] <- rf$rsq[length(rf$rsq)]
  IncMSE[i,3] <- mean(MSE[,1])
  IncMSE[i,4] <- best[i]
  imp <- importance(rf)
  for (j in 1:nrow(imp)){
    IncMSE[i,j+4] <- imp[j,1]
    IncNodePurity[i,j+1] <- imp[j,2]
  }
}
```


## Outputs

### Importance plot
```{r}
varImpPlot(rf)
```

### Increased MSE table
```{r}
library(knitr)
library(kableExtra)
library(dplyr)

t <- IncMSE %>% mutate_if(is.numeric, format, digits=2, nsmall=0) %>% t(.)
for (i in 1:ncol(t)){
  colnames(t)[i] <- t[1,i]
}
t <- t[-1,]

  t %>% kbl() %>% kable_paper(full_width=F) %>%
  column_spec (1:ncol(t),border_left = T, border_right = T, width="1cm") %>%
  kable_styling()
```

### Normalized increased MSE table
```{r}
IncMSE_norm <- IncMSE
range = vector()
IncMSE_norm[IncMSE_norm<0] <- 0

for (i in 5:ncol(IncMSE_norm)){
  min <- min(IncMSE_norm[,i])
  max <- max(IncMSE_norm[,i])
  range[i] <- max - min
  for (j in 1:nrow(IncMSE_norm)){
    IncMSE_norm[j,i] <- (IncMSE_norm[j,i]-min)/range[i]
  }
}

t <- IncMSE_norm %>% mutate_if(is.numeric, format, digits=2, nsmall=0) %>% t(.)
for (i in 1:ncol(t)){
  colnames(t)[i] <- t[1,i]
}
t <- t[-1,]

  t %>% kbl() %>% kable_paper(full_width=F) %>%
  column_spec (1:ncol(t),border_left = T, border_right = T, width="1cm") %>%
  kable_styling()
```

### Increased Node Purity
```{r}
library(knitr)
library(kableExtra)
library(dplyr)

t <- IncNodePurity %>% t(.)
for (i in 1:ncol(t)){
  colnames(t)[i] <- t[1,i]
}
t <- t[-1,]

  t %>% kbl() %>% kable_paper(full_width=F) %>%
  column_spec (1:ncol(t),border_left = T, border_right = T, width="1cm") %>%
  kable_styling()
```

### Normalized Increased Node Purity table
```{r}
IncMSE_norm <- IncNodePurity
range = vector()
IncMSE_norm[IncMSE_norm<0] <- 0

for (i in 2:ncol(IncMSE_norm)){
  min <- min(IncMSE_norm[,i])
  max <- max(IncMSE_norm[,i])
  range[i] <- max - min
  for (j in 1:nrow(IncMSE_norm)){
    IncMSE_norm[j,i] <- (IncMSE_norm[j,i]-min)/range[i]
  }
}

t <- IncMSE_norm %>% mutate_if(is.numeric, format, digits=2, nsmall=0) %>% t(.)
for (i in 1:ncol(t)){
  colnames(t)[i] <- t[1,i]
}
t <- t[-1,]

  t %>% kbl() %>% kable_paper(full_width=F) %>%
  column_spec (1:ncol(t),border_left = T, border_right = T, width="1cm") %>%
  kable_styling()
```